---
title: "CatchEfficiency"
author: "Tobias Schwoerer"
date: "May 18, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

Fishing vessels in Alaska fall into two categories for registration purposes: 1) Coast Guard certified, and 2) not CG certified. For 1) the owners needs to register with the CG, for 2) registration with DMV occurs. Each boat has a unique ADFG ID that stays with the boat over its lifetime (valid for skiffs and boats in the case of Purse Seine)

#Importing vessel information from CFEC
```{r setup, include=FALSE}
library(Hmisc)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(data.table)
library(RColorBrewer)
library(Hmisc)

#importing datasets on fishing vessels available from CFEC and the registered boats database from DMV 
vessels <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/knb.92147.1",stringsAsFactors = FALSE)
 
regions <-  read.csv("E:/GoogleDrive/CURRENT_PROJECTS/3_SASAP_salmon_synthesis/Analysis_ref_files/index_table_051118.csv", stringsAsFactors=FALSE)

#calculating age of vessels
cfec <- mutate(vessels, age = Year-Year.Built)
#flagging Coast Guard registered vessels
cfec$CG <- ifelse(cfec$Coast.Guard..=="",0,1)
#flagging salmon boats
cfec$salmon <- ifelse(cfec$Salmon.Registration.Area=="N/A",0,1)

#subsets of salmon boats
salmon_fleet <- subset(cfec, salmon==1)
salmon_fleet$SASAP.Region <- with(salmon_fleet,ifelse(Salmon.Registration.Area=="Yakutat","Southeast",ifelse(Salmon.Registration.Area=="Peninsula/Aleutian Islands","Alaska Peninsula and Aleutian Islands", ifelse(Salmon.Registration.Area=="Lower Yukon","Yukon", Salmon.Registration.Area))))


#coding residency status to match other datasets coding convention
salmon_fleet$Residency <- ifelse(salmon_fleet$State =="AK","res","non")
#participating in other than salmon fisheries, as indicated by other non-salmon gear
salmon_fleet$other.gear <- with(salmon_fleet, ifelse(Freezer.Canner=="Yes"|Tender.Packer=="Yes"|Charter=="Yes"|Long.Line=="Yes"|Otter.Trawl=="Yes"|Pots=="Yes"|Beam.Trawl=="Yes"|Scallop.Dredge=="Yes"|Mechanical.Jig=="Yes"|Double.Otter.Trawl=="Yes"|Herring.Gill.Net=="Yes"|Pair.Trawl=="Yes"|Diving.Hand.Picking=="Yes",1,0))

#flagging Copper Riverdrift gillnetters and adding SASAP.Region Copper River
salmon_fleet$SASAP.Region <- with(salmon_fleet, ifelse(SASAP.Region=="Prince William Sound" & Drift.Gill.Net =="Yes","Copper River",SASAP.Region))

#recoding yes values to numeric 1 and 0 for gear
salmon_fleet <- salmon_fleet%>%
  mutate_at(vars(29:49),funs(ifelse(Purse.Seine=="Yes"|Beach.Seine=="Yes"|Drift.Gill.Net=="Yes"|Set.Gill.Net=="Yes"|Hand.Troll=="Yes"|Fish.Wheel=="Yes"|Power.Troll=="Yes",1,0)))
salmon_fleet$Refrigeration <- ifelse(salmon_fleet$Refrigeration=="Yes",1,0)
```


#Creating summary tables for fleet
```{r}
#salmon_gear <- c("Purse.Seine","Beach.Seine","Drift.Gill.Net","Set.Gill.Net","Hand.Troll","Fish.Wheel","Power.Troll")

fleet_sum <- salmon_fleet %>%
  gather(Gear,cases, Purse.Seine,Drift.Gill.Net,Beach.Seine,Set.Gill.Net,Hand.Troll,Fish.Wheel,Power.Troll)%>%
  filter(cases==1)%>%
  group_by(SASAP.Region,Year,Gear,Residency)%>%
  summarise(AvgHP = mean(Horse.Power),
            AvgTons = mean(Gross.Tons),
            AvgAge = mean(age),
            fleetSize = n_distinct(ADFG.Number))

```
What I learned:
Looks like the maximum horsepower data are outliers or crappy data. Rather than taking the sum of horsepower across the fleet, maybe better to stick to the mean and compare to mean catch.


#Importing BIT harvest and exvessel data from CFEC
```{r}
BIT <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Aaa308395-54f6-412c-9ca3-0112a1d67938", stringsAsFactors=FALSE)
#deleting the space in the fisheries code to be consistent with other datasets
BIT$Fishery <- gsub(" ", "", BIT$Fishery, fixed = TRUE)
#salmon permit subset
salmon <- unique(BIT$Fishery[grep("^S", BIT$Fishery)])
BITs <- subset(BIT, Fishery %in% salmon)

#determining fisheries and years for which we have average earnings (pounds), then fill that in for any missing and throw out remaining missing (19 records with fewer than 3 permits fished)
BITs$Resident.Average.Earnings <- ifelse(is.na(BITs$Resident.Average.Earnings), BITs$Average.Earnings,BITs$Resident.Average.Earnings)
BITs$Resident.Average.Pounds <- ifelse(is.na(BITs$Resident.Average.Pounds), BITs$Average.Pounds,BITs$Resident.Average.Pounds)
BITs$Nonresident.Average.Earnings <- ifelse(is.na(BITs$Nonresident.Average.Earnings), BITs$Average.Earnings,BITs$Nonresident.Average.Earnings)
BITs$Nonresident.Average.Pounds <- ifelse(is.na(BITs$Nonresident.Average.Pounds), BITs$Average.Pounds,BITs$Nonresident.Average.Pounds)

BITs <- BITs %>% 
  drop_na(Resident.Average.Earnings,Nonresident.Average.Earnings,Resident.Average.Pounds,Nonresident.Average.Pounds)

#join SASAP.Regions
fishery_region <- read.csv("E:/GoogleDrive/CURRENT_PROJECTS/3_SASAP_salmon_synthesis/Analysis_ref_files/fishery_codes_regions.csv", stringsAsFactors=FALSE)

BITs <- BITs %>%
  left_join(fishery_region, by=c("Fishery"))

#joining consistent gear types with the vessel data
gearCodes <- read.csv("E:/GoogleDrive/CURRENT_PROJECTS/3_SASAP_salmon_synthesis/Analysis_ref_files/gear_codes_vessels.csv",stringsAsFactors=F)
BITs$GearCode <- as.numeric(substr(BITs$Fishery, 2, 3))
BITs <- BITs%>%
  left_join(gearCodes,by=c("GearCode"="CODE"))

#Creating summary tables
BIT_sum <- BITs%>%
  gather(Residency,HarvestLbs,Resident.Average.Pounds, Nonresident.Average.Pounds)
BIT_sum$Residency <- ifelse(BIT_sum$Residency=="Resident.Average.Pounds","res","non")

BIT_sum2 <- BIT_sum%>%
  group_by(SASAPRegions,Year, ADFG_H_GEAR_CODE,Residency)%>%
  summarise(MeanLbs = mean(HarvestLbs))
#Note, for the most part, the mean remains the average harvest in line above
```

#Combining harvest data
```{r}
fleet <- fleet_sum%>%
 left_join(BIT_sum2, by=c("SASAP.Region"="SASAPRegions","Year","Residency","Gear"="ADFG_H_GEAR_CODE"))%>%
 filter(MeanLbs!=0)%>%
 drop_na(MeanLbs)

#Creating subset for Purse Seine and Drift Gill Net
PSGN <- subset(fleet, Gear=="Purse.Seine" |Gear=="Drift.Gill.Net")
PSGN$eff <- PSGN$AvgHP/PSGN$MeanLbs*1000 
#eliminating outliers
PSGN <- subset(PSGN, eff<=60)

PS <- subset(PSGN, Gear=="Purse.Seine" )
DG <- subset(PSGN, Gear=="Drift.Gill.Net" )
```

#Plotting mean Horse Power and HP/1000lbs of fish caught
```{r}

ggplot(PSGN) + 
  facet_wrap(~SASAP.Region,ncol = 3, scales = "fixed")+
  geom_line(aes(x=Year,y=AvgHP,color=Gear,linetype= Residency), size=1, show.legend=T, text = element_text(size=30)) + 
  labs(title="Mean Horse Power of Salmon Fishing Fleet by Region, Gear, and Owner Residence (1978-2016)",x="Year",y="mean HP") +
  theme_bw()





ggplot(BIT_sum2) + facet_wrap(~SASAPRegions,ncol = 3, scales = "free")+ labs(title="Average harvest (1975-2016)",x="Year",y="mean lbs") + geom_point(aes(x=Year,y=MeanLbs/1000,color=Residency),size=1) +  geom_smooth(aes(x=Year, y=MeanLbs/1000, color=Residency),span=1) + theme_bw() + opts(legend.position = "bottom")


ggplot(DG) + facet_wrap(~SASAP.Region,ncol = 3, scales = "free")+ labs(title="Drift Gillnet Horse Power per 1000lbs Caught by Region and Gear (1975-2016)",x="Year",y="mean HP/ mean catch in 1000's lbs") + geom_point(aes(x=Year,y=eff,color=Gear),size=1, show.legend=F,text = element_text(size=20)) +geom_smooth(aes(x=Year, y=eff, color=Gear),span=0.7,show.legend=F) +ylim(0,20)+ theme_bw(base_size = 15)

ggplot(PS) + facet_wrap(~SASAP.Region,ncol = 3, scales = "free")+ labs(title="Drift Gillnet Horse Power per 1000lbs Caught by Region and Gear (1975-2016)",x="Year",y="mean HP/ mean catch in 1000's lbs") + geom_point(aes(x=Year,y=eff,color=blues9),size=1, show.legend=F,text = element_text(size=20)) +geom_smooth(aes(x=Year, y=eff, color=blues9),span=0.7,show.legend=F) +ylim(0,20)+ theme_bw(base_size = 15)


```









Importing ADFG harvest dataset (NOTE this is not the BIT from CFEC), adding SASAP regions (excludes EEZ catch), had to import by hand due to commas being part of netpounds turning those into factors instead of numeric variables
Datasets available at https://knb.ecoinformatics.org/#view/urn:uuid:40473bde-9774-4581-aafb-5d2c3b4a70d1
```{r}
#Importing area by YEAR dataset
h <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A49a6d2fa-d47a-482c-8a28-d8a9623dc16e", stringsAsFactors = FALSE)

#Importing district by YEAR dataset
DbyY <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A6619074a-3037-41a6-8dbf-9bf2988cabec", stringsAsFactors = FALSE)

#Taking out Kotzebue,Norton Sound, replacing it with district wide data from the DbyY dataset. Note, the h dataset which is AreaByYEAR will contain data that may be confidential in the DistrictByYEAR dataset, so going with AreaByYEAR first is better. Also excluding EEZ harvest
keep <- c("Bristol Bay","Chignik","Cook Inlet Lower","Cook Inlet Upper","Juneau Hanies","Ketchikan Craig","Kodiak","Kuskokwim","North Peninsula","PWS Copper Bering Rivers","PWS Eastern District","PWS excluding CRBRED","Petersburg Wrangell","Sitka Pelican","South Peninsula","Yakutat","Yukon","Dutch Harbor","Adak Western Aleutians")
h_ext <- subset (h, AREA %in% keep)  
#drop the Confidential column
h_ext <- subset(h_ext, select = -CONFIDENTIAL)

#Adding SASAP regions
harvest_regions <- read.csv("E:/GoogleDrive/CURRENT_PROJECTS/3_SASAP_salmon_synthesis/Analysis_ref_files/harvest_regions.csv", stringsAsFactors=FALSE)
h_ext2 <- h_ext %>% 
  full_join(harvest_regions, by = c("AREA" = "ADFG_Area")) %>%
  na.omit()
h_ext2 <- subset(h_ext2, select = -AREA)
names(h_ext2)[14]<-"SASAP.Region"
h_ext2 <-h_ext2 [c("SASAP.Region","YEAR","ADFG_H_GEAR_CODE","ADFG_I_SPECIES_CODE","No_of_Fish","Value","NETPOUNDS","WholeLbs","people_count","processor_count","vessel_count","Gear","Species")]

#Aggregating district by YEAR data to SASAP region level
district_region_lut <- read.csv("E:/GoogleDrive/CURRENT_PROJECTS/3_SASAP_salmon_synthesis/Analysis_ref_files/district_region_lut.csv", stringsAsFactors=FALSE)
lut <- district_region_lut
DbyY <- left_join(DbyY, lut)

dist_agg <- DbyY %>% 
  group_by(SASAP.Region,YEAR,ADFG_H_GEAR_CODE, ADFG_I_SPECIES_CODE, Gear,Species) %>% 
  summarise_at(vars(Value:vessel_count), sum, na.rm = TRUE)%>%
  drop_na(Value)

#ordering columns for later combination with the area by YEAR data
dist_agg <-dist_agg [c("SASAP.Region","YEAR","ADFG_H_GEAR_CODE","ADFG_I_SPECIES_CODE","No_of_Fish","Value","NETPOUNDS","WholeLbs","people_count","processor_count","vessel_count","Gear","Species")]

#Subsetting for Norton Sound and Kotzebue
norton <- subset(dist_agg,SASAP.Region=="Norton Sound")
kotz <- subset(dist_agg,SASAP.Region=="Kotzebue")

#Combining the above to have one dataset for SASAP regions
harvest_comb <- bind_rows(h_ext2,kotz,norton)

#Cleaning trivial data, e.g. gear=herring gillnet, and removing odd records and gear types with very few and small records e.g. pelagic trawl
harvest_comb <- subset(harvest_comb, Gear !="pelagic/mid-water trawl")
harvest_comb <- subset(harvest_comb, Gear !="herring gillnet")
```


