---
title: "Harvest"
author: "Tobias Schwoerer"
date: "March 19, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---
Importing dataset, adding SASAP regions (excludes EEZ catch), had to import by hand due to commas being part of netpounds turning those into factors instead of numeric variables
Datasets available at https://knb.ecoinformatics.org/#view/urn:uuid:40473bde-9774-4581-aafb-5d2c3b4a70d1
```{r}
library(dplyr)
library(ggplot2)
library(data.table)
library(tidyr)
library(RColorBrewer)

#Importing area by YEAR dataset
h <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A49a6d2fa-d47a-482c-8a28-d8a9623dc16e", stringsAsFactors = FALSE)

#Importing district by YEAR dataset
DbyY <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A6619074a-3037-41a6-8dbf-9bf2988cabec", stringsAsFactors = FALSE)

```

Taking out Kotzebue,Norton Sound, replacing it with district wide data from the DbyY dataset. Note, the h dataset which is AreaByYEAR will contain data that may be confidential in the DistrictByYEAR dataset, so going with AreaByYEAR first is better. Also excluding EEZ harvest
```{r}
keep <- c("Bristol Bay","Chignik","Cook Inlet Lower","Cook Inlet Upper","Juneau Hanies","Ketchikan Craig","Kodiak","Kuskokwim","North Peninsula","PWS Copper Bering Rivers","PWS Eastern District","PWS excluding CRBRED","Petersburg Wrangell","Sitka Pelican","South Peninsula","Yakutat","Yukon","Dutch Harbor","Adak Western Aleutians")
h_ext <- subset (h, AREA %in% keep)  
#drop the Confidential column
h_ext <- subset(h_ext, select = -CONFIDENTIAL)

#Adding SASAP regions
harvest_regions <- read.csv("E:/GoogleDrive/CURRENT_PROJECTS/3_SASAP_salmon_synthesis/Analysis_ref_files/harvest_regions.csv", stringsAsFactors=FALSE)
h_ext2 <- h_ext %>% 
  full_join(harvest_regions, by = c("AREA" = "ADFG_Area")) %>%
  na.omit()
h_ext2 <- subset(h_ext2, select = -AREA)
names(h_ext2)[14]<-"SASAP.Region"
h_ext2 <-h_ext2 [c("SASAP.Region","YEAR","ADFG_H_GEAR_CODE","ADFG_I_SPECIES_CODE","No_of_Fish","Value","NETPOUNDS","WholeLbs","people_count","processor_count","vessel_count","Gear","Species")]
```

Aggregating district by YEAR data to SASAP region level
```{r}
district_region_lut <- read.csv("E:/GoogleDrive/CURRENT_PROJECTS/3_SASAP_salmon_synthesis/Analysis_ref_files/district_region_lut.csv", stringsAsFactors=FALSE)
lut <- district_region_lut
DbyY <- left_join(DbyY, lut)

dist_agg <- DbyY %>% 
  group_by(SASAP.Region,YEAR,ADFG_H_GEAR_CODE, ADFG_I_SPECIES_CODE, Gear,Species) %>% 
  summarise_at(vars(Value:vessel_count), sum, na.rm = TRUE)%>%
  drop_na(Value)

#ordering columns for later combination with the area by YEAR data
dist_agg <-dist_agg [c("SASAP.Region","YEAR","ADFG_H_GEAR_CODE","ADFG_I_SPECIES_CODE","No_of_Fish","Value","NETPOUNDS","WholeLbs","people_count","processor_count","vessel_count","Gear","Species")]

#Subsetting for Norton Sound and Kotzebue
norton <- subset(dist_agg,SASAP.Region=="Norton Sound")
kotz <- subset(dist_agg,SASAP.Region=="Kotzebue")
```

Combining the above to have one dataset for SASAP regions
```{r}
harvest_comb <- bind_rows(h_ext2,kotz,norton)
```

Cleaning trivial data, e.g. gear=herring gillnet, and removing odd records and gear types with very few and small records e.g. pelagic trawl
```{r}
harvest_comb <- subset(harvest_comb, Gear !="pelagic/mid-water trawl")
harvest_comb <- subset(harvest_comb, Gear !="herring gillnet")
```

Calculating a mean weight by species, year, and region. This is then used to estimate harvest weight for sport and subsistence fisheries
```{r}
a1 <- harvest_comb
a1 <- subset(a1,No_of_Fish!=0)
a <- a1 %>%
  mutate(avg_wgt=NETPOUNDS/No_of_Fish)%>%
  group_by(Species, SASAP.Region, YEAR)%>%
  summarise(median_wgt = quantile(avg_wgt, probs = c(0.5)),
                   p05_wgt = quantile(avg_wgt, probs = c(0.05)),
                   p95_wgt = quantile(avg_wgt, probs = c(0.95))) 
a_state <- a1 %>%
  mutate(avg_wgt=NETPOUNDS/No_of_Fish)%>%
  group_by(Species, YEAR)%>%
  summarise(median_wgt = quantile(avg_wgt, probs = c(0.5)),
                   p05_wgt = quantile(avg_wgt, probs = c(0.05)),
                   p95_wgt = quantile(avg_wgt, probs = c(0.95))) 


chinook <- subset(a,Species=="chinook")
chinook <- ggplot(chinook, aes(x=YEAR,y=median_wgt)) + facet_wrap(~SASAP.Region,ncol = 3, scales = "fixed")+ labs(title="Median, 5th,and 95th percentile in weight of commercially caught chinook salmon by region",x="Year",y="lbs per fish") + geom_ribbon(aes(ymin = p05_wgt, ymax = p95_wgt), fill = "grey70") + geom_line(aes(y = median_wgt))
chinook

all_other <- subset(a,Species!="chinook")
#cleaning out data for pink in Yukon that looks odd
all_other <- subset(all_other, !(Species=="pink"&SASAP.Region=="Yukon"))

sockeye <- subset(all_other,Species=="sockeye")
sockeye <- ggplot(sockeye, aes(x=YEAR,y=median_wgt)) + facet_wrap(~SASAP.Region,ncol = 3, scales = "fixed")+ labs(title="Median, 5th, and 95th percentile of sockeye weight by region",x="Year",y="lbs per fish") + geom_ribbon(aes(ymin = p05_wgt, ymax = p95_wgt), fill = "grey70") + geom_line(aes(y = median_wgt))
sockeye

coho <- subset(all_other,Species=="coho")
coho <- ggplot(coho, aes(x=YEAR,y=median_wgt)) + facet_wrap(~SASAP.Region,ncol = 3, scales = "fixed")+ labs(title="Median, 5th, and 95th percentile of coho weight by region",x="Year",y="lbs per fish") + geom_ribbon(aes(ymin = p05_wgt, ymax = p95_wgt), fill = "grey70") + geom_line(aes(y = median_wgt))
coho

chum <- subset(all_other,Species=="chum")
chum <- ggplot(chum, aes(x=YEAR,y=median_wgt)) + facet_wrap(~SASAP.Region,ncol = 3, scales = "fixed")+ labs(title="Median, 5th, and 95th percentile of chum weight by region",x="Year",y="lbs per fish") + geom_ribbon(aes(ymin = p05_wgt, ymax = p95_wgt), fill = "grey70") + geom_line(aes(y = median_wgt))
chum

pink <- subset(all_other,Species=="pink")
pink <- ggplot(pink, aes(x=YEAR,y=median_wgt)) + facet_wrap(~SASAP.Region,ncol = 3, scales = "fixed")+ labs(title="Median, 5th, and 95th percentile of pink weight by region",x="Year",y="lbs per fish") + geom_ribbon(aes(ymin = p05_wgt, ymax = p95_wgt), fill = "grey70") + geom_line(aes(y = median_wgt))
pink

statewide_weights <- ggplot(a_state, aes(x=YEAR,y=median_wgt)) + facet_wrap(~Species,ncol = 3, scales = "free") + labs(title="Statewide Median, 5th, and 95th percentile weight of commercially caught salmon by species",x="Year",y="lbs per fish")+ geom_ribbon(aes(ymin = p05_wgt, ymax = p95_wgt), fill = "grey70") + geom_line(aes(y = median_wgt))
statewide_weights
  
```

Filling zeros for missing data in certain YEARs or for certain species ( better for graphing later)
```{r}
h1 <- harvest_comb
h2 <-expand.grid(SASAP.Region=unique(h1$SASAP.Region), YEAR=1975:2016, Gear=unique(h1$Gear), Species=unique(h1$Species))
h3 <- merge(h2,h1, by.x=c("SASAP.Region", "YEAR", "Gear","Species"),by.y=c("SASAP.Region", "YEAR", "Gear","Species"), all.x = TRUE)
h3[is.na(h3)]<-0

harvest<-h3
```


Fraction of commercial catch by geartype and species statewide
```{r}
h_gear_species <- harvest %>%
    group_by(Species, Gear, YEAR) %>%
    dplyr::summarise(sum_lbs = sum(NETPOUNDS), 
                     sum_value = sum(Value),
                     sum_fish = sum(No_of_Fish))

h_total_species <- h_gear_species %>%
    group_by(Species, YEAR) %>%
    dplyr::summarise(total_lbs = sum(sum_lbs), 
                     total_value = sum(sum_value), 
                     total_fish = sum(sum_fish))

h_prop_species <- h_gear_species %>% 
  full_join(h_total_species, by = c("Species", "YEAR")) %>%
  na.omit() %>%
  dplyr::mutate(prop_lbs = sum_lbs/total_lbs,
                prop_value = sum_value/total_value, 
                prop_no_fish = sum_fish/total_fish) 
```

Figures for proportion of total harvest weight caught by gear type by species
```{r}
ggplot(h_prop_species, aes(x=YEAR,y=prop_lbs, fill=Gear)) + geom_area(alpha=0.6, size=1,colour="black") + facet_wrap(~Species,ncol = 3, scales = "free")+ labs(fill = "Gear",title="Statewide Commercial Salmon Harvest by Species and Gear Type",x="Year",y="Proportion of harvest weight")
```


Fraction of commercial catch by gear type by region over time
##Needs to be fixed to exclude some of the rare gear types
```{r}
h_sum_by_gear <- harvest %>%
    group_by(SASAP.Region, Gear, YEAR) %>%
    dplyr::summarise(sum_lbs = sum(NETPOUNDS), 
                     sum_value = sum(Value),
                     sum_fish = sum(No_of_Fish))

h_total <- h_sum_by_gear %>%
    group_by(SASAP.Region, YEAR) %>%
    dplyr::summarise(total_lbs = sum(sum_lbs), 
                     total_value = sum(sum_value), 
                     total_fish = sum(sum_fish))

h_prop <- h_sum_by_gear %>% 
  full_join(h_total, by = c("SASAP.Region", "YEAR")) %>%
  na.omit() %>%
  dplyr::mutate(prop_lbs = sum_lbs/total_lbs,
                prop_value = sum_value/total_value, 
                prop_no_fish = sum_fish/total_fish) 
```

Figures for proportion of total harvest weight caught by gear type
```{r}
ggplot(h_prop, aes(x=YEAR,y=prop_lbs, fill=Gear)) + geom_area(alpha=0.6, size=1,colour="black") + facet_wrap(~SASAP.Region,ncol = 3, scales = "free") + labs(fill = "Gear",title="Commercial Salmon Harvest by Region and Gear Type",x="Year",y="Proportion of harvest weight")
```

Fraction of commercial catch by species, gear type, and region over time
```{r}
h_sum_by_gearS <- harvest %>%
    group_by(SASAP.Region, Gear, YEAR, Species) %>%
    dplyr::summarise(sum_lbs = sum(NETPOUNDS), 
                     sum_value = sum(Value),
                     sum_fish = sum(No_of_Fish))

h_totalS <- h_sum_by_gearS %>%
    group_by(SASAP.Region, YEAR, Species) %>%
    dplyr::summarise(total_lbs = sum(sum_lbs), 
                     total_value = sum(sum_value), 
                     total_fish = sum(sum_fish))

h_propS <- h_sum_by_gearS %>% 
  full_join(h_totalS, by = c("SASAP.Region", "YEAR","Species")) %>%
  na.omit() %>%
  dplyr::mutate(prop_lbs = sum_lbs/total_lbs,
                prop_value = sum_value/total_value, 
                prop_no_fish = sum_fish/total_fish) 


```

Figures for proportion of total harvest weight by species, gear type, and region over time
```{r}
chinook <- ggplot(subset(h_propS,Species=="chinook"), aes(x=YEAR,y=prop_lbs, fill=Gear)) + geom_area(alpha=0.6, size=1,colour="black") + facet_wrap(~SASAP.Region,ncol = 3, scales = "fixed")+ labs(fill = "Gear",title="Commercial Chinook Harvest by Region and Gear Type",x="Year", y="Proportion of harvest weight")
chinook

chum <- ggplot(subset(h_propS,Species=="chum"), aes(x=YEAR,y=prop_lbs, fill=Gear)) + geom_area(alpha=0.6, size=1,colour="black") + facet_wrap(~SASAP.Region,ncol = 3, scales = "fixed")+ labs(fill = "Gear",title="Commercial Chum Harvest by Region and Gear Type",x="Year", y="Proportion of harvest weight")
chum

coho <- ggplot(subset(h_propS,Species=="coho"), aes(x=YEAR,y=prop_lbs, fill=Gear)) + geom_area(alpha=0.6, size=1,colour="black") + facet_wrap(~SASAP.Region,ncol = 3, scales = "fixed")+ labs(fill = "Gear",title="Commercial Coho Harvest by Region and Gear Type",x="Year", y="Proportion of harvest weight")
coho

pink <- ggplot(subset(h_propS,Species=="pink"), aes(x=YEAR,y=prop_lbs, fill=Gear)) + geom_area(alpha=0.6, size=1,colour="black") + facet_wrap(~SASAP.Region,ncol = 3, scales = "fixed") + labs(fill = "Gear", title="Commercial Pink Harvest by Region and Gear Type",x="Year", y="Proportion of harvest weight")
pink

sockeye <- ggplot(subset(h_propS,Species=="sockeye"), aes(x=YEAR,y=prop_lbs, fill=Gear)) + geom_area(alpha=0.6, size=1,colour="black") + facet_wrap(~SASAP.Region,ncol = 3, scales = "fixed")+ labs(fill = "Gear", title="Commercial Sockeye Harvest by Region and Gear Type",x="Year", y="Proportion of harvest weight")
sockeye
```


# Sport fish harvest available at: https://knb.ecoinformatics.org/#view/urn:uuid:6a6a530f-3660-424f-adab-c771d1c89a5d
```{r}
sf_h <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A22ef5022-8ade-4549-acac-c18656dd2033", stringsAsFactors = FALSE)


sf_c <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A542141d3-ed5a-4d97-b759-28a17757b0b8", stringsAsFactors = FALSE)


#renaming Alaska Peninsula Region name
sf_h2 <- sf_h %>%
     mutate(region_name=replace(region_name,  region_name=="Aleutian Islands", "Alaska Peninsula and Aleutian Islands")) 

#add weight information from commercial fisheries catch to sport fish harvest, for missing values use statewide mean weight by species
sf_h2 <- sf_h2 %>% 
  full_join(a, by = c("region_name" = "SASAP.Region","species"="Species", "year"="YEAR"))

sf_h3 <- sf_h2 %>%
    left_join(a_state, by = c("species"="Species","year"="YEAR"))

sf_h3$median_wgt.x <- ifelse(is.na(sf_h3$median_wgt.x),sf_h3$median_wgt.y,sf_h3$median_wgt.x)
sf_h3$p05_wgt.x <- ifelse(is.na(sf_h3$p05_wgt.x),sf_h3$p05_wgt.y,sf_h3$p05_wgt.x)
sf_h3$p95_wgt.x <- ifelse(is.na(sf_h3$p95_wgt.x),sf_h3$p95_wgt.y,sf_h3$p95_wgt.x)
sf_h3 <- subset(sf_h3, select = -c(p95_wgt.y, p05_wgt.y,median_wgt.y))
 #renaming 
names(sf_h3)[names(sf_h3)=="median_wgt.x"] <- "median_wgt"
names(sf_h3)[names(sf_h3)=="p05_wgt.x"] <- "p05_wgt"
names(sf_h3)[names(sf_h3)=="p95_wgt.x"] <- "p95_wgt"
```

Calculating sport fish harvest weight for resident and non-resident point estimates, the total harvest weights, also rearrange dataset
```{r}
sfHarvest <- filter(sf_h3,sf_h3$TYPE=="0-EST")
sfHarvest$harvest_lbs <- sfHarvest$median_wgt*sfHarvest$harvest

#combining non-resident and resident sport fish harvest and effort
sfHarTot <- sfHarvest%>%
  group_by(year, region_name, species)%>%
  summarise(total_lbs = sum(harvest_lbs),
            total_DF = sum(DF),
            total_FISH = sum(FISH_TOTAL),
            total_salmon = sum(harvest))
sfHarTot$salmon_score <- sfHarTot$total_salmon/sfHarTot$total_FISH
```

Figures for proportion of salmon caught to overall sport fish caught fish by species and region
```{r}
ggplot(sfHarTot, aes(x=year,y=salmon_score, fill=species)) + geom_area(alpha=0.6, size=1,colour="black") + facet_wrap(~region_name,ncol = 3, scales = "fixed")+ labs(fill = "Species",title="Proportion of salmon harvested among all sportfish by region and species",x="Year",y="Proportion of harvested fish")
```

#Subsistence harvest available at https://knb.ecoinformatics.org/#view/urn:uuid:a7d46a96-2bbf-4131-a291-339b950b334e , this file includes only "All Permit Types" therefore, we need to subtract the personal use harvest and add it later, recoding the "permit type" 
```{r}
ss <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A552cfff5-c192-411a-a9ab-11ffe434dbc4", stringsAsFactors = FALSE)

ss <- mutate_at(ss,vars(Species), funs(tolower))

ss2 <- ss%>%                
  group_by(SASAP.Region, Community, Year, PermitTypeName, Species)%>%
  summarise(HarvestLbs = sum(HarvestLbs))

pu <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A638adfdc-8313-45f5-b437-886556dfdf34", stringsAsFactors = FALSE)
PUregions <- read.csv("E:/GoogleDrive/CURRENT_PROJECTS/3_SASAP_salmon_synthesis/Analysis_ref_files/pu_regions.csv", stringsAsFactors = FALSE) 

pu <- pu %>%
  left_join(PUregions, by="FisheryName")

pu <- mutate_at(pu,vars(Species), funs(tolower))
pu2 <- pu%>%                
  group_by(SASAP.Region, Community, Year, PermitTypeName, Species)%>%
  summarise(HarvestLbs = sum(HarvestLbs))

pu_ss_comb <- ss2 %>%
  full_join(pu2, by=c("SASAP.Region","Community","Year","Species"))%>%
  mutate(HarvestLbs = HarvestLbs.x-HarvestLbs.y)%>%
  drop_na(SASAP.Region)

#renaming some columns, and recoding SASAP.Regions to be consistent
names(pu_ss_comb)[names(pu_ss_comb)=="HarvestLbs.y"] <- "PU"
names(pu_ss_comb)[names(pu_ss_comb)=="HarvestLbs"] <- "SS"
pu_ss_comb$SASAP.Region <- ifelse(pu_ss_comb$SASAP.Region=="Aleutian Islands","Alaska Peninsula and Aleutian Islands",pu_ss_comb$SASAP.Region)

pu_ss_comb$SS <- ifelse(is.na(pu_ss_comb$SS), pu_ss_comb$HarvestLbs.x, pu_ss_comb$SS)
pu_ss_comb$PU <- ifelse(is.na(pu_ss_comb$PU), 0, pu_ss_comb$PU)

pu_ss_comb2 <-pu_ss_comb %>% 
   gather(PermitType,lbs, SS,PU)
#cleaning up columns
ss3 <- subset(pu_ss_comb2, select = -c(PermitTypeName.x,HarvestLbs.x,PermitTypeName.y))
# aggregating by region to take out by community detail (community is the city of residence of the subsistence permit owner)
ss4 <- ss3%>%
  group_by(SASAP.Region,Year,PermitType,Species)%>%
  summarise(harvestLbs=sum(lbs))
ss4[is.na(ss4)]<-0

subsistence <- subset(ss4,PermitType=="SS",select = -PermitType)
personalUse <- subset(ss4,PermitType=="PU", select = -PermitType)
```

Combining datasets
```{r}
#combining gear types in commercial data
harvest2 <- harvest%>%
  group_by(Species, SASAP.Region, YEAR)%>%
  summarise(No_of_Fish = sum(No_of_Fish),
            NETPOUNDS = sum(NETPOUNDS))
#combining datasets
comb_sf_com <- harvest2%>%
  dplyr::full_join(sfHarTot, by = c("SASAP.Region"="region_name","Species"="species","YEAR"="year"))  
comb_sf_com[is.na(comb_sf_com)]<-0

comb_sf_ss_com <- comb_sf_com %>%
  dplyr::full_join(subsistence, by = c("SASAP.Region","Species","YEAR"="Year"))%>%
  dplyr::full_join(personalUse, by = c("SASAP.Region","Species","YEAR"="Year"))
  
comb_sf_ss_com[is.na(comb_sf_ss_com)]<-0
  
  

#summarizing all salmon species
sf_com <- comb_sf_ss_com%>%
  group_by(SASAP.Region, YEAR)%>%
  summarise(sportfish = sum(total_lbs)/1000000,
            commercial = sum(NETPOUNDS)/1000000,
            subsistence = sum(harvestLbs.x)/1000000,
            personalUse = sum(harvestLbs.y)/1000000)

sf_com_raw <- comb_sf_ss_com%>%
  group_by(SASAP.Region, YEAR)%>%
  summarise(sportfish = sum(total_lbs),
            commercial = sum(NETPOUNDS),
            subsistence = sum(harvestLbs.x),
            personalUse = sum(harvestLbs.y))


#subsetting to only years where all sportfish, commercial, and subsistence data is available, which is 1996 on
plot_data <- subset(sf_com, YEAR>=1996)
#gathering for plotting
plot_data <- plot_data%>%
  gather(stakeholder,million_lbs, commercial,sportfish, subsistence, personalUse)

CI_plot <- subset(plot_data, SASAP.Region =="Cook Inlet")
CR_plot <- subset(plot_data, SASAP.Region =="Copper River")
```

Harvest lbs by stakeholder group by region
```{r}
ggplot(plot_data, aes(x=YEAR,y=million_lbs,fill=stakeholder)) + geom_bar(stat="identity",position=position_stack(reverse=T)) + facet_wrap(~SASAP.Region,ncol = 3, scales = "free")+ labs(title="Salmon harvested by stakeholder group by region",x="Year",y="million lbs harvested") +theme(legend.position = c(0.9, 0.05)) + scale_fill_brewer(palette="Spectral")

CI <- ggplot(CI_plot, aes(x=YEAR,y=million_lbs,fill=stakeholder)) + geom_bar(stat = "identity", position=position_stack(reverse=T)) + labs(title="Salmon harvested by stakeholder group in Cook Inlet",x="Year",y="million lbs harvested") + scale_fill_brewer(palette="Spectral")
CI

CR <- ggplot(CR_plot, aes(x=YEAR,y=million_lbs,fill=stakeholder)) + geom_bar(stat = "identity", position=position_stack(reverse=T)) + labs(title="Salmon harvested by stakeholder group in the Copper River Region",x="Year",y="million lbs harvested") + scale_fill_brewer(palette="Spectral")
CR

```

Statewide harvest lbs by stakeholder allocation
```{r}
sf_com_state <- comb_sf_ss_com%>%
  group_by(YEAR)%>%
  summarise(sportfish = sum(total_lbs)/1000000,
            commercial = sum(NETPOUNDS)/1000000,
            subsistence = sum(HarvestLbs)/1000000,
            total = sum(sportfish,commercial,subsistence))


#subsetting to only years where all sportfish, commercial, and subsistence data is available, which is 1996 on
plot_data_state <- subset(sf_com_state, YEAR>=2015)   
#gathering for plotting
plot_data_state <- plot_data_state%>%
  gather(stakeholder,million_lbs, commercial,sportfish, subsistence)
```


COPIED FROM ELSEWHERE
Fraction of commercial catch by species, gear type, and region over time
```{r}
h_sum_by_gearS <- harvest %>%
    group_by(SASAP.Region, Gear, YEAR, Species) %>%
    dplyr::summarise(sum_lbs = sum(NETPOUNDS), 
                     sum_value = sum(Value),
                     sum_fish = sum(No_of_Fish))

h_totalS <- h_sum_by_gearS %>%
    group_by(SASAP.Region, YEAR, Species) %>%
    dplyr::summarise(total_lbs = sum(sum_lbs), 
                     total_value = sum(sum_value), 
                     total_fish = sum(sum_fish))

h_propS <- h_sum_by_gearS %>% 
  full_join(h_totalS, by = c("SASAP.Region", "YEAR","Species")) %>%
  na.omit() %>%
  dplyr::mutate(prop_lbs = sum_lbs/total_lbs,
                prop_value = sum_value/total_value, 
                prop_no_fish = sum_fish/total_fish) 


```

Figures for proportion of total harvest weight by species, gear type, and region over time
```{r}
chinook <- ggplot(subset(h_propS,Species=="chinook"), aes(x=YEAR,y=prop_lbs, fill=Gear)) + geom_area(alpha=0.6, size=1,colour="black") + facet_wrap(~SASAP.Region,ncol = 3, scales = "fixed")+ labs(fill = "Gear",title="Commercial Chinook Harvest by Region and Gear Type",x="Year", y="Proportion of harvest weight")
chinook

```





